Constructing Sailing Match Race Schedules: Round Robin Pairing Lists

### 2015-04-07 paper.tex
Unused
LaTeX tweaks

\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
\usepackage{url}
\usepackage{graphicx}
\usepackage{float}
\usepackage{listings}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathtools}
\usepackage[rightcaption]{sidecap}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage{framed}
%\usepackage{coffee2}
\usepackage{tikz}

\usepackage{color}
\newcommand{\todo}[1]{\textcolor{blue}{#1}}

@@

This work describes a round robin competition format that arises from sailing, known as match
racing. Competitors, i.e. {\em skippers}, compete against each other in a series of matches taking
place in rounds, known as {\em flights}.  A match is composed of two skippers, with each skipper in
a boat on his own. Skippers in a match set off together, one on the port side, the other starboard,
and first home wins that match. Criteria published by the International Sailing Federation (ISAF)
dictate what makes a legal match race schedule and what should be optimised. This is a rich set of
criteria, 13 in all, and as far as we know, all published schedules have been produced by hand. This
is a daunting task. A close inspection of these schedules reveals that most are illegal, violating
many of the match race scheduling criteria, many schedules are missing and some that are legal are
far from optimal.

This paper describes the scheduling problem. We present a constraint programming solution and
document how this model was constructed incrementally. We believe this illustrates how adaptable
constraint programming is, i.e. given a fairly robust model this can be incrementally developed to
address all of the required criteria. Our schedules are then constructed in a sequence of stages,
not dissimilar to the hybrid and decomposition based approaches in \cite{lombardi2012}. Some ISAF
published schedules are presented along with some of our own, some of these being new.  These
schedules are essentially blank forms, made available to event organisers, which can be filled in
with names of actual competitors, i.e.\ these schedules are reusable.

@@

Next, we note that skippers can be ranked prior to an event -- based on their performance in
previous events\footnote{\url{http://www.sailing.org/rankings/match/}} -- which is used to seed the
skippers. The ordering of skippers within a match signifies their starting position (port or
starboard) -- as skippers allocated the starboard starting position gain a small competitive
advantage in that match - criteria 8 accomplishes the seedings.

Criteria 10 is ambiguous: - this could be interpreted on a schedule basis: i.e.\ to minimise the
overall number of changes in the entire pairing list schedule; or alternatively as well as a fair,
but minimal number of changes across all skippers. In our work, we take the latter viewpoint.

@@

In Figure~\ref{oriented}, port is red and starboard green. The sequence of zero/one variables for
each skipper is shown on the left. Bottom right is the DFA, and top right is the final schedule
produced from stages 1 to 4. The schedule has 6 boat changes and a maximum imbalance of 1. Three
days CPU time was devoted to stages 1 and 2. Stages 3 and 4 completed in 15 milliseconds. The
schedule presented here is new, not appearing in the example schedules provided in \cite{isaf}.%ISAF
World Sailing Umpire's Manual

@@

We present three pairs of schedules. The first pair is for 8 skippers and 6 boats, one published by
ISAF, the other generated by us. Next, we present two new schedules, not appearing in the ISAF
manual, for 9 skippers and 6 boats and 9 skippers with 8 boats. The last pair is for 10 skippers and
6 boats, again ISAF published and generated by us. For each of our schedules, we used 1 day cpu time
for the first stage (minimising boat changes) then a further 2 days to minimise imbalance. Stages 3
and 4 (renaming and orientation) completed in less than a second.

\begin{center}
\begin{figure}[h]
\begin{minipage}[t]{0.5\textwidth}
\hspace{0.6cm}
%\begin{scriptsize}
\begin{tabular}{cccc}
        \toprule
        Flight & \multicolumn{3}{c}{Matches} \\ \midrule
        0 & (5,2) & (4,3) & (1,6) \\
        1 & (4,2)  & (6,5) & (3,1) \\
        2 & (6,4) & (2,1) & (3,5) \\
        3 & (6,2) & (5,0) & (1,7) \\
        4 & (5,1) & (0,6) & (2,7) \\
        5 & (7,5) & (2,0) & (6,3) \\
        6 & (0,7) & (4,1) & (3,2) \\
       7 & (7,4)  & (0,3) &  \\
       8 & (4,0)  & (7,3) &  \\
       9 & (5,4)  & (7,6) & (1,0) \\ \bottomrule
    \end{tabular}
%\end{scriptsize}
\label{08-06a}
\end{minipage}
\hfill
\begin{minipage}[t]{0.5\textwidth}
\hspace{0.6cm}
%\begin{scriptsize}
\begin{tabular}{cccc}
        \toprule
        Flight & \multicolumn{3}{c}{Matches} \\ \midrule
        0 & (2,7) & (3,0) & (5,4) \\
        1 & (0,2)  & (4,3) & (7,5) \\
        2 & (0,4) & (7,6) & (5,3) \\
        3 & (4,6) & (5,0) & (3,1) \\
        4 & (6,5) & (3,2) & (1,4) \\
        5 & (6,3) & (4,7) & (2,1) \\
        6 & (7,3) & (1,6) & (4,2) \\
       7 & (7,1)  & (2,5) & (6,0) \\
       8 & (5,1)  & (0,7) & (6,2) \\
       9 & (1,0)  & & \\ \bottomrule
    \end{tabular}
%\end{scriptsize}
\label{08-06b}
\end{minipage}
\caption{Schedules for 8 skippers and 6 boats. On the left, the published ISAF schedule (illegal)
and on the right is our schedule.}
\label{08-06}
\end{figure}
\end{center}

\noindent
\subsection{8 skippers and 6 boats}

We first analyse the ISAF schedule, Table \ref{08-06} on the
left. Criteria 4 (skippers in last match in a flight cannot be first in next flight) is violated on
3 occasions (skipper 1 in flight 3, skipper 7 in flight 4, skipper 0 in flight 7). Criteria 5 (no
more than two consecutive port or starboard assignments) is violated for skipper 6 (flights 1 to 3)
and 7 (flights 7 to 9). Criteria 6 (imbalance) is violated for skippers 6 and 8. Criteria 12  (new
boats do not sail first) is violated in flight 9 for skipper 5. Finally, the schedule has 8 boat
changes. Our schedule, Table \ref{08-06} on the right, respects all criteria and has 6 boat changes
and a maximum imbalance of 1.

\subsection{9 skippers}

Two new schedules are presented in Table \ref{09}, on the left for 6 boats
and on the right 8 boats. Neither of these schedules appear in the ISAF Umpires' Manual. Both
schedules respect all criteria. For 6 boats there are 8 boat changes (no skipper making more than 1
boat change) and a maximum imbalance of 2. For 8 boats there are again 8 boat changes, no skipper
with more than 1 boat change, and perfect balance (each skipper 4 times on starboard side and 4 on
port side).

\begin{center}
\begin{figure}[h]
\begin{minipage}[t]{0.5\textwidth}
\hspace{0.6cm}
%\begin{scriptsize}
\begin{tabular}{cccc}
        \toprule
        Flight & \multicolumn{3}{c}{Matches} \\ \midrule
        0 & (0,7) & (3,2) & (5,4) \\
        1 & (0,2) & (7,4) & (5,3) \\
        2 & (4,0) & (7,3) & (2,5) \\
        3 & (3,0) & (2,4) & (6,5) \\
        4 & (4,3) & (5,1) & (8,6) \\
        5 & (3,1) & (6,2) & (4,8) \\
        6 & (6,3) & (1,4) & (8,2) \\
        7 & (4,6) & (2,7) & (1,8) \\
        8 & (7,6) & (0,8) & (2,1) \\
        9 & (8,7) & (5,0) & (6,1) \\
        10 & (8,5) & (1,7) & (0,6) \\
        11 & (7,5) & (3,8) & (1,0) \\ \bottomrule
    \end{tabular}
%\end{scriptsize}
\label{09-06}
\end{minipage}
%\hfill
\begin{minipage}[t]{0.5\textwidth}
\hspace{0.6cm}
%\begin{scriptsize}
\begin{tabular}{ccccc}
        \toprule
        Flight & \multicolumn{4}{c}{Matches} \\ \midrule
        0 & (0,6) & (3,2) & (5,4) & (1,7) \\
        1 & (2,0) & (6,3) & (4,1) & (8,7) \\
        2 & (6,4) & (0,5) & (7,2) & (3,8) \\
        3 & (4,7) & (1,8) & (0,3) & (5,2) \\
        4 & (3,1) & (7,5) & (8,0) & (2,6) \\
        5 & (3,7) & (4,8) & (2,1) & (6,5) \\
        6 & (8,2) & (1,6) & (5,3) & (0,4) \\
        7 & (5,1) & (7,0) & (6,8) & (4,3) \\
        8 & (8,5) & (2,4) & (7,6) & (1,0) \\ \bottomrule
        & & & & \\
        & & & & \\
        & & & & \\
    \end{tabular}
%\end{scriptsize}
\label{09-08}
\end{minipage}
\caption{Two new schedules. On the left, 9 skippers and 6 boats and on the right 9 skippers and 8 boats.}
\label{09}
\end{figure}
\end{center}

\subsection{10 skippers and 6 boats}

Table \ref{10-06} shows the ISAF schedule on the right for 10 skippers and 6 boats and on the right,
our schedule.  The ISAF schedule violates criteria 5 (no more than 2 consecutive port or starboard
assignments) 10 times (twice for skipper 0, 1 and 6 and once for skippers 3, 7, 8 and 9).  Criteria
12 (new boats do not sail first) is violated 7 times (in flight 3 for skippers 6 and 2, flight 5 for
skipper 8, flight 6 for skipper 7, flight 8 for skipper 1, flight 9 for skipper 4 and flight 12 for
skipper 4 again). There are 22 boat changes, with the minimum changes for a skipper being 1 and the
maximum 3. Our schedule, right of Table \ref{10-06} satisfies all criteria, has 12 boat changes with
the minimum changes for a skipper being 0 and the maximum 2, and a maximum imbalance of 1.

\begin{center}
\begin{figure}[h]
\begin{minipage}[t]{0.5\textwidth}
\hspace{0.6cm}
%\begin{scriptsize}
\begin{tabular}{cccc}
 \toprule
Flight & \multicolumn{3}{c}{Matches} \\ \midrule
0 & (8,3) &  (1,7) &  (0,9) \\
1 & (7,3) &  (0,8) &  (1,9) \\
2 & (0,7) &  (9,3) &  (1,8) \\
3 & (6,2) &  (5,0) &  (7,4) \\
4 & (5,2) &  (6,4) &  (3,0) \\
5 & (8,6) &  (2,9) &  (4,1) \\
6 & (7,2) &  (4,8) &  (9,6) \\
7 & (2,4) &  (0,6) &  (7,5) \\
8 & (6,1) &  (5,9) &  (8,2) \\
9 & (9,4) &  (8,7) &  (6,5) \\
10 & (9,7) &  (6,3) &  (8,5) \\
11 & (7,6) &  (3,1) &  (9,8) \\
12 & (4,3) &  (5,1) &  (2,0) \\
13 & (3,5) &  (4,0) &  (2,1) \\
14 & (5,4) &  (3,2) &  (1,0) \\ \bottomrule
\end{tabular}
%\end{scriptsize}
\label{10-06a}
\end{minipage}
%\hfill
\begin{minipage}[t]{0.5\textwidth}
\hspace{0.6cm}
%\begin{scriptsize}
\begin{tabular}{ccccc}
\toprule
Flight & \multicolumn{3}{c}{Matches} \\ \midrule
0 & (4,7) & (3,2) & (0,5) \\
1 & (2,4) & (7,0) & (3,5) \\
2 & (0,4) & (7,3) & (5,2) \\
3 & (4,3) & (0,2) & (5,7) \\
4 & (3,0) & (6,5) & (2,7) \\
5 & (6,3) & (7,1) & (8,2) \\
6 & (7,6) & (2,1) & (9,8) \\
7 & (2,6) & (8,7) & (4,9) \\
8 & (6,8) & (1,4) & (7,9) \\
9 & (8,1) & (5,4) & (9,6) \\
10 & (5,1) & (9,2) & (4,6) \\
11 & (1,9) & (6,0) & (8,4) \\
12 & (0,9) & (3,8) & (6,1) \\
13 & (8,0) & (9,5) & (1,3) \\
14 & (5,8) & (9,3) & (1,0) \\ \bottomrule
\end{tabular}
%\end{scriptsize}
\label{10-06b}
\end{minipage}
\caption{Schedules for 10 skippers and 6 boats. On the left, the ISAF published schedule (illegal) and on the right, our schedule.}
\label{10-06}
\end{figure}
\end{center}

@@

\subsection{Discussion}

It was anticipated that this problem would be easy to model. This naivety was due to "round robin",
leading us to believe that we could use some of the models proposed in \cite{trick2002},
\cite{trick2008} and \cite{henz2004}. This naivety slowly dissolved as we addressed more and more of
the ISAF criteria. A number of models we produced were scrapped due to misunderstanding about the
actual problem, i.e. a communication problem between the authors. Eventually this was resolved by
presenting the majority of the criteria as a deterministic finite automaton. This became our Rosetta
Stone, with the surprising benefit that it not only improved the teams communications, it was also a
constraint!

@@

We used the choco constraint programming toolkit and one of our goals was to use only toolkit
constraints, i.e. we wanted to see how far we could go without implementing our own specialised
constraints. We are pleased to say that we did not have to do anything out of the box, although we
did feel that we had used just about every constraint there was.

There are many redundant variables in our model. One of our first valid models was essentially that
shown in Figure \ref{skipper5}, and did not use $modMatch$ and $match$ variables. Performance was
hopeless, struggling to produce a 7 skipper 6 boat schedule in days. The $modMatch$ and $match$
variables were added. This improved domain filtering, and a flattened $match$ array was used as
decision variables. That is, the decisions where ``when do we schedule this match?".  At this point
we had not yet produced a 9 skipper 6 boat schedule and we suspected that the combined criteria
might exclude such a schedule. A non-constraint model was developed by the fourth author, primarily
to prove that there was no schedule for 9 skippers with 6 boats. This program used a backtracking
search with decision variables being positions within flights, i.e. time slots being assigned
matches. Prior to search, symmetry breaking was performed, and at the top of search the matches in
the first flight were set. A solution was found in seconds! With this experience we added in the
$time$ variables, channeled these into the existing model and used these as decision variables, i.e.
the question now was ``what match will we do now?". We also anchored the matches in the first
flight. With these changes we began to produce schedules in acceptable time. The model was then
built upon, incrementally adding criteria. This took a surprisingly short amount of time, sometimes
minutes of coding to add in a new feature. The model was then enhanced so that it would address
optimisation rather than just satisfaction, again a trivial programming task.

@@

So, why did we use constraint programming? The answer is obvious:  we are constraint programmers, or
to borrow Mark Twain's words ``To a man with a hammer, everything looks like a nail." But actually,
constraint programming has been a good way to go. From an engineering perspective, it has allowed us
to prototype solutions quickly and to build solutions incrementally. There is also an advantage that
we might exploit in the future: we can now investigate the effect different criteria have on our
ability to produce schedules and how relaxation of those affect optimisation criteria. That is, the
constraint program might be used to design the next set of criteria for match race schedules, or
allow the event organiser to decide which criteria he may sacrifice to ensure a faster schedule with
less boat changes.

>>>>

\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
\usepackage{url}
\usepackage{graphicx}
\usepackage{float}
\usepackage{listings}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathtools}
\usepackage[rightcaption]{sidecap}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage{framed}
\usepackage{tikz}

@@

This work describes a round robin competition format that arises from sailing, known as match
racing. Competitors, i.e.\ \emph{skippers}, compete against each other in a series of matches taking
place in rounds, known as \emph{flights}.  A match is composed of two skippers, with each skipper in
a boat on his own. Skippers in a match set off together, one on the port side, the other starboard,
and first home wins that match. Criteria published by the International Sailing Federation (ISAF)
dictate what makes a legal match race schedule and what should be optimised. This is a rich set of
criteria, 13 in all, and as far as we know, all published schedules have been produced by hand. This
is a daunting task. A close inspection of these schedules reveals that most are illegal, violating
many of the match race scheduling criteria, many schedules are missing, and some that are legal are
far from optimal.

This paper describes the scheduling problem. We present a constraint programming solution and
document how this model was constructed incrementally. We believe this illustrates how adaptable
constraint programming is, i.e.\ given a fairly robust model this can be incrementally developed to
address all of the required criteria. Our schedules are then constructed in a sequence of stages,
not dissimilar to the hybrid and decomposition based approaches in \cite{lombardi2012}. Some ISAF
published schedules are presented along with some of our own, some of these being new.  These
schedules are essentially blank forms, made available to event organisers, which can be filled in
with names of actual competitors, i.e.\ these schedules are reusable.

@@

Next, we note that skippers can be ranked prior to an event, based on their performance in
previous events\footnote{\url{http://www.sailing.org/rankings/match/}}, which is used to seed the
skippers. The ordering of skippers within a match signifies their starting position (port or
starboard), as skippers allocated the starboard starting position gain a small competitive
advantage in that match---criteria 8 accomplishes the seedings.

Criteria 10 is ambiguous: this could be interpreted on a schedule basis: i.e.\ to minimise the
overall number of changes in the entire pairing list schedule; or alternatively as well as a fair,
but minimal number of changes across all skippers. In our work, we take the latter viewpoint.

@@

In Figure~\ref{oriented}, port is red and starboard green. The sequence of zero/one variables for
each skipper is shown on the left. Bottom right is the DFA, and top right is the final schedule
produced from stages 1 to 4. The schedule has 6 boat changes and a maximum imbalance of 1. Three
days CPU time was devoted to stages 1 and 2. Stages 3 and 4 completed in 15 milliseconds. The
schedule presented here is new, not appearing in the example schedules provided in \cite{isaf}.

@@

We present three pairs of schedules. The first pair is for 8 skippers and 6 boats, one published by
ISAF, the other generated by us. Next, we present two new schedules, not appearing in the ISAF
manual, for 9 skippers and 6 boats and 9 skippers with 8 boats. The last pair is for 10 skippers and
6 boats, again ISAF published and generated by us. For each of our schedules, we used 1 day CPU time
for the first stage (minimising boat changes) then a further 2 days to minimise imbalance. Stages 3
and 4 (renaming and orientation) completed in less than a second.

\begin{center}
\begin{figure}[h]
\begin{minipage}[t]{0.5\textwidth}
    \centering
\begin{tabular}{cccc}
        \toprule
        Flight & \multicolumn{3}{c}{Matches} \\ \midrule
        0 & (5,2) & (4,3) & (1,6) \\
        1 & (4,2)  & (6,5) & (3,1) \\
        2 & (6,4) & (2,1) & (3,5) \\
        3 & (6,2) & (5,0) & (1,7) \\
        4 & (5,1) & (0,6) & (2,7) \\
        5 & (7,5) & (2,0) & (6,3) \\
        6 & (0,7) & (4,1) & (3,2) \\
       7 & (7,4)  & (0,3) &  \\
       8 & (4,0)  & (7,3) &  \\
       9 & (5,4)  & (7,6) & (1,0) \\ \bottomrule
    \end{tabular}
\label{08-06a}
\end{minipage}
\hfill
\begin{minipage}[t]{0.5\textwidth}
    \centering
\begin{tabular}{cccc}
        \toprule
        Flight & \multicolumn{3}{c}{Matches} \\ \midrule
        0 & (2,7) & (3,0) & (5,4) \\
        1 & (0,2)  & (4,3) & (7,5) \\
        2 & (0,4) & (7,6) & (5,3) \\
        3 & (4,6) & (5,0) & (3,1) \\
        4 & (6,5) & (3,2) & (1,4) \\
        5 & (6,3) & (4,7) & (2,1) \\
        6 & (7,3) & (1,6) & (4,2) \\
       7 & (7,1)  & (2,5) & (6,0) \\
       8 & (5,1)  & (0,7) & (6,2) \\
       9 & (1,0)  & & \\ \bottomrule
    \end{tabular}
\label{08-06b}
\end{minipage}
\caption{Schedules for 8 skippers and 6 boats. On the left, the published ISAF schedule (illegal),
and on the right is our schedule.}
\label{08-06}
\end{figure}
\end{center}

\paragraph{8 skippers and 6 boats:} We first analyse the ISAF schedule, Table~\ref{08-06} on the
left. Criteria 4 (skippers in last match in a flight cannot be first in next flight) is violated on
3 occasions (skipper 1 in flight 3, skipper 7 in flight 4, skipper 0 in flight 7). Criteria 5 (no
more than two consecutive port or starboard assignments) is violated for skipper 6 (flights 1 to 3)
and 7 (flights 7 to 9). Criteria 6 (imbalance) is violated for skippers 6 and 8. Criteria 12  (new
boats do not sail first) is violated in flight 9 for skipper 5. Finally, the schedule has 8 boat
changes. Our schedule, Table~\ref{08-06} on the right, respects all criteria and has 6 boat changes
and a maximum imbalance of 1.

\paragraph{9 skippers}: Two new schedules are presented in Table~\ref{09}, on the left for 6 boats
and on the right 8 boats. Neither of these schedules appear in the ISAF Umpires' Manual. Both
schedules respect all criteria. For 6 boats there are 8 boat changes (no skipper making more than 1
boat change) and a maximum imbalance of 2. For 8 boats there are again 8 boat changes, no skipper
with more than 1 boat change, and perfect balance (each skipper 4 times on starboard side and 4 on
port side).

\begin{center}
\begin{figure}[h]
\begin{minipage}[t]{0.5\textwidth}
\centering
\begin{tabular}{cccc}
        \toprule
        Flight & \multicolumn{3}{c}{Matches} \\ \midrule
        0 & (0,7) & (3,2) & (5,4) \\
        1 & (0,2) & (7,4) & (5,3) \\
        2 & (4,0) & (7,3) & (2,5) \\
        3 & (3,0) & (2,4) & (6,5) \\
        4 & (4,3) & (5,1) & (8,6) \\
        5 & (3,1) & (6,2) & (4,8) \\
        6 & (6,3) & (1,4) & (8,2) \\
        7 & (4,6) & (2,7) & (1,8) \\
        8 & (7,6) & (0,8) & (2,1) \\
        9 & (8,7) & (5,0) & (6,1) \\
        10 & (8,5) & (1,7) & (0,6) \\
        11 & (7,5) & (3,8) & (1,0) \\ \bottomrule
    \end{tabular}
\centering
\label{09-06}
\end{minipage}
\hfill
\begin{minipage}[t]{0.5\textwidth}
\centering
\begin{tabular}{ccccc}
        \toprule
        Flight & \multicolumn{4}{c}{Matches} \\ \midrule
        0 & (0,6) & (3,2) & (5,4) & (1,7) \\
        1 & (2,0) & (6,3) & (4,1) & (8,7) \\
        2 & (6,4) & (0,5) & (7,2) & (3,8) \\
        3 & (4,7) & (1,8) & (0,3) & (5,2) \\
        4 & (3,1) & (7,5) & (8,0) & (2,6) \\
        5 & (3,7) & (4,8) & (2,1) & (6,5) \\
        6 & (8,2) & (1,6) & (5,3) & (0,4) \\
        7 & (5,1) & (7,0) & (6,8) & (4,3) \\
        8 & (8,5) & (2,4) & (7,6) & (1,0) \\ \bottomrule
        & & & & \\
        & & & & \\
        & & & & \\
    \end{tabular}
\label{09-08}
\end{minipage}
\caption{Two new schedules. On the left, 9 skippers and 6 boats, and on the right 9 skippers and 8 boats.}
\label{09}
\end{figure}
\end{center}

\paragraph{10 skippers and 6 boats:} Table \ref{10-06} shows the ISAF schedule on the right for 10
skippers and 6 boats and on the right, our schedule.  The ISAF schedule violates criteria 5 (no more
than 2 consecutive port or starboard assignments) 10 times (twice for skipper 0, 1 and 6 and once
for skippers 3, 7, 8 and 9).  Criteria 12 (new boats do not sail first) is violated 7 times (in
flight 3 for skippers 6 and 2, flight 5 for skipper 8, flight 6 for skipper 7, flight 8 for skipper
1, flight 9 for skipper 4 and flight 12 for skipper 4 again). There are 22 boat changes, with the
minimum changes for a skipper being 1 and the maximum 3. Our schedule, right of Table \ref{10-06}
satisfies all criteria, has 12 boat changes with the minimum changes for a skipper being 0 and the
maximum 2, and a maximum imbalance of 1.

\begin{center}
\begin{figure}[h]
\begin{minipage}[t]{0.5\textwidth}
    \centering
\begin{tabular}{cccc}
 \toprule
Flight & \multicolumn{3}{c}{Matches} \\ \midrule
0 & (8,3) &  (1,7) &  (0,9) \\
1 & (7,3) &  (0,8) &  (1,9) \\
2 & (0,7) &  (9,3) &  (1,8) \\
3 & (6,2) &  (5,0) &  (7,4) \\
4 & (5,2) &  (6,4) &  (3,0) \\
5 & (8,6) &  (2,9) &  (4,1) \\
6 & (7,2) &  (4,8) &  (9,6) \\
7 & (2,4) &  (0,6) &  (7,5) \\
8 & (6,1) &  (5,9) &  (8,2) \\
9 & (9,4) &  (8,7) &  (6,5) \\
10 & (9,7) &  (6,3) &  (8,5) \\
11 & (7,6) &  (3,1) &  (9,8) \\
12 & (4,3) &  (5,1) &  (2,0) \\
13 & (3,5) &  (4,0) &  (2,1) \\
14 & (5,4) &  (3,2) &  (1,0) \\ \bottomrule
\end{tabular}
\label{10-06a}
\end{minipage}
\hfill
\begin{minipage}[t]{0.5\textwidth}
\centering
\begin{tabular}{ccccc}
\toprule
Flight & \multicolumn{3}{c}{Matches} \\ \midrule
0 & (4,7) & (3,2) & (0,5) \\
1 & (2,4) & (7,0) & (3,5) \\
2 & (0,4) & (7,3) & (5,2) \\
3 & (4,3) & (0,2) & (5,7) \\
4 & (3,0) & (6,5) & (2,7) \\
5 & (6,3) & (7,1) & (8,2) \\
6 & (7,6) & (2,1) & (9,8) \\
7 & (2,6) & (8,7) & (4,9) \\
8 & (6,8) & (1,4) & (7,9) \\
9 & (8,1) & (5,4) & (9,6) \\
10 & (5,1) & (9,2) & (4,6) \\
11 & (1,9) & (6,0) & (8,4) \\
12 & (0,9) & (3,8) & (6,1) \\
13 & (8,0) & (9,5) & (1,3) \\
14 & (5,8) & (9,3) & (1,0) \\ \bottomrule
\end{tabular}
\label{10-06b}
\end{minipage}
\caption{Schedules for 10 skippers and 6 boats. On the left, the ISAF published schedule (illegal),
and on the right, our schedule.}
\label{10-06}
\end{figure}
\end{center}

@@

\section{Discussion}

It was anticipated that this problem would be easy to model. This naivety was due to ``round
robin'', leading us to believe that we could use some of the models proposed in \cite{trick2002},
\cite{trick2008} and \cite{henz2004}. This naivety slowly dissolved as we addressed more and more of
the ISAF criteria. A number of models we produced were scrapped due to misunderstanding about the
actual problem, i.e.\ a communication problem between the authors. Eventually this was resolved by
presenting the majority of the criteria as a deterministic finite automaton. This became our Rosetta
Stone, with the surprising benefit that it not only improved the teams communications, it was also a
constraint!

@@

We used the choco constraint programming toolkit and one of our goals was to use only toolkit
constraints, i.e.\ we wanted to see how far we could go without implementing our own specialised
constraints. We are pleased to say that we did not have to do anything out of the box, although we
did feel that we had used just about every constraint there was.

There are many redundant variables in our model. One of our first valid models was essentially that
shown in Figure~\ref{skipper5}, and did not use $\modMatch$ and $\match$ variables. Performance was
hopeless, struggling to produce a 7 skipper 6 boat schedule in days. The $\modMatch$ and $\match$
variables were added. This improved domain filtering, and a flattened $\match$ array was used as
decision variables. That is, the decisions where ``when do we schedule this match?''.  At this point
we had not yet produced a 9 skipper 6 boat schedule and we suspected that the combined criteria
might exclude such a schedule. A non-constraint model was developed by the fourth author, primarily
to prove that there was no schedule for 9 skippers with 6 boats. This program used a backtracking
search with decision variables being positions within flights, i.e.\ time slots being assigned
matches. Prior to search, symmetry breaking was performed, and at the top of search the matches in
the first flight were set. A solution was found in seconds! With this experience we added in the
$time$ variables, channeled these into the existing model and used these as decision variables,
i.e.\ the question now was ``what match will we do now?''. We also anchored the matches in the first
flight. With these changes we began to produce schedules in acceptable time. The model was then
built upon, incrementally adding criteria. This took a surprisingly short amount of time, sometimes
minutes of coding to add in a new feature. The model was then enhanced so that it would address
optimisation rather than just satisfaction, again a trivial programming task.

@@

So, why did we use constraint programming? The answer is obvious:  we are constraint programmers, or
to borrow Mark Twain's words ``To a man with a hammer, everything looks like a nail''. But actually,
constraint programming has been a good way to go. From an engineering perspective, it has allowed us
to prototype solutions quickly and to build solutions incrementally. There is also an advantage that
we might exploit in the future: we can now investigate the effect different criteria have on our
ability to produce schedules and how relaxation of those affect optimisation criteria. That is, the
constraint program might be used to design the next set of criteria for match race schedules, or
allow the event organiser to decide which criteria he may sacrifice to ensure a faster schedule with
less boat changes.
